\documentclass{article}
\usepackage[UTF8, scheme = plain, zihao = 5, linespread = 1.3]{ctex}
\usepackage{indentfirst}
\input{up2020.cfg}
\newcommand*{\cf}{\emph{cf.}}
\newcommand*{\eg}{\emph{eg.}}
\newcommand*{\vs}{\emph{vs.}}
\newcommand*{\etc}{\emph{etc}}
\newcommand*{\etal}{\emph{et al}}
\newcommand*{\rnrs}[1]{\emph{R$^{#1}\!$RS}}
\newcommand*{\colskipa}{\vspace{-0.5\baselineskip}}
\newcommand*{\colskipb}{\vspace{-0.5\baselineskip}}

\begin{document}
\title{\textbf{Unix Philosophy 2020}}
\author{%
	Casper Ti.\ Vector
	\texttt{<\url{CasperVector@gmail.com}>}\\
	\url{https://gitea.com/CasperVector/up2020}%
}
\date{2019-08/2019-09; version \docversion}
\maketitle
\vspace{\baselineskip}
\tableofcontents

\newpartx{0.5}
\specialsec{Foreword}

I began learning Linux in the end of 2008, and using it in the beginning
of 2009; I have always used Linux as my main operating system thereafter.
During the course of using Linux and interacting with the ecosystem of
Unix-like operating systems, I got attracted to the notion of Unix
philosophy, and have benefited enormously from it.  (For more details on
the origin of this document, see the \hyperref[sec:afterword]{Afterword}.)
This document is a compendium of my thoughts about the Unix philosophy,
organised into three main parts, as are summarised below.

In the first part, we will first discuss the question ``\stress{what is the
essence of the Unix philosophy}?'', and I will argue that it is the minimisation
of the cognitive complexity of the system while almost satisfying the
requirements.  These years, many people think the philosophy, which originated
from legacy limitations on computational resources back in the 1970s, is no
longer applicable in software development.  So \stress{is the Unix philosophy
still relevant}, and will it still be relevant in the 2020s?  My answer is
``yes, and more relevant than when it was born''.  As a conclusion to this part,
I will give my experience about \stress{how the Unix philosophy can be followed
in actual use and development}, to the advantage of the user~/ developer.

In the second part, we will move our focus out of the field of programming, and
explore \stress{the applicability of the Unix philosophy in science, technology
and society, and the limits to that applicability}.  After the discussion on
science and technology, I will consider minimalism in philosophy, literature
and arts, and attempt to find a cognitive basis for minimalism and its limits;
the theory on the cognitive limits to minimalism will then be used to explain
why society often does not work like Unix.  As a conclusion, we will discuss
\stress{how minimalism can be adhered to, especially by the ordinary person},
to boost one's efficiency in everyday work.

In the third part, we will return to programming, but this time we will begin
with a digression about Lisp and especially Scheme, which also encourage
minimalism.  In order to understand the minimalism of Lisp, we will examine
the notion of homoiconicity, which I find as important as complexity.  I will
then try to answer ``\stress{what is the root of the antagonism between
Unix and Lisp}?'', and finally explore the benefits and feasibility of
\stress{a minimalist language framework which adopts the best
from Lisp and C}, and which also avoids the weaknesses.

Before actually diving into the main parts, it should be noted that this
document is meant to be read critically, for two reasons.  First, what I write
here is just my personal opinions about the Unix philosophy, and the latter
is just one way, among the many ones, to model the world.  Second, I still
know little about unfamiliar research fields like philosophy, cognitive
science, and programming languages, so the contents about these topics are
highly tentative, and there may well be points that turn out to be amateurish
or even ludicrous.  I attempted to avoid making mistakes in this document,
and please feel free to tell me if you find any.

This document is an extended transcript of my technical report,
\emph{Unix philosophy in the contemporary era}\cupercite{casper2018}, on
10 November 2018 for the Linux Club of Peking University; I consider it
as the culmination of my 10-year experience in the open-source community,
and dedicate it to the 50th anniversary of Unix\cupercite{salus2005}.
I enjoyed making the report as well as this document, and wish you
could also have some fun reading through the document, whether you
are a newcomer to the community, or an expert almost 3 times my age.

\newpage
\specialsec{How to read this document}

Due to feedback from multiple readers, it has been brought to my attention that
the nature of this document as a compendium -- a collection of thoughts about
various subjects -- can lead to some confusions: who are expected to read this
document, and how should they read the document?  Here I will briefly explain
the logical relation between the parts~/ sections in this document, and then
attempt to give a guide to readers with various backgrounds; please feel free
to tell me if you have any advice on how to make this document more accessible.

\begin{wquoting}
\begin{tikzpicture}[
	every node/.style = {draw, rectangle, font = \linespread{1}\selectfont},
	every matrix/.style = {
		draw = none, anchor = west, row sep = 1em, column sep = 2em
	}, ll/.style = {align = left}, rr/.style = {align = right}
]
	\newcommand*{\subj}[4]{\node [#1] {\textbf{#2:} #3\\#4}}
	\matrix {
		\subj{rr}{01--07}{modernising}{the Unix philosophy}; &[0.4em]
		\subj{ll}{13--18}{minimalism}{outside of programming}; &
		\subj{ll}{22--24}{Unix}{philosophy and Lisp}; \\
		\subj{rr}{08--10}{significance}{of the philosophy}; &
		\subj{rr}{19}{minimalism from}{a cognitive viewpoint}; &
		\subj{rr}{25--27}{reconciling}{Unix and Lisp}; \\
		\subj{rr}{11--12}{real-world}{Unix philosophy}; &
		\subj{ll}{20--21}{minimalism for}{society and the individual}; \\
	};
\end{tikzpicture}
\end{wquoting}

Above is an outline of the logical organisation of this document.  In my
opinion, nowadays the Unix philosophy is often neglected, questioned or
even distorted, due to misunderstanding about it.  So in the first part,
I will first present my attempt at a modernised formulation of the
philosophy in Sections~\ref{sec:intro}--\ref{sec:complex}.  Using the
formulation as the theoretical foundation, I will discuss significance
of the Unix philosophy in the contemporary context in Sections~%
\ref{sec:quality}--\ref{sec:foss}, and real-world applications
of the philosophy in Sections~\ref{sec:devel}--\ref{sec:user}.

I disagree with some people's opinion that the Unix philosophy was now
merely an aesthetic, because I believe it is rooted in human recognition,
which results in its inherent importance.  In the second part, using the
above-mentioned formulation, I will first examine some analogues to the Unix
philosophy outside of the field of programming in Sections~\ref{sec:hilbert}%
--\ref{sec:art}.  After that, I will present my cognitive justification
for the philosophy in Section~\ref{sec:cognitive}, and discuss its
value for society and the individual in Sections~\ref{sec:society}--%
\ref{sec:worklife}, similar to the last two sections in the first part.

In the recent years, it has gradually become my opinion that expertise from
the Unix and Lisp circles can be joined to build systems that are more
Unix-ish than was possible before.  So in the third part, I will first
discuss the relation between Lisp and the previously formulated Unix
philosophy in Sections~\ref{sec:lisp}--\ref{sec:wib}.  Finally, I will
examine the benefits of the proposed Unix~/ Lisp reconciliation, and how this
reconciliation can be achieved in Sections~\ref{sec:benefits}--\ref{sec:toe}.

It is generally recommended to read this document in the original order, and
just skip parts~/ sections you are less interested in; however, Section~%
\ref{sec:complex} lays the theoretical foundation for this document, so you
should at least skim through it.  Quite a few examples from the ``init war''
are used in the first part, so people entangled in related events are advised
to read that part.  Those who like philosophical discussions might feel
interested in the second part.  Unix developers interested in programming
languages, as well as researchers of these languages, will perhaps find
the third part attractive.  Footnotes are generally less essential
contents, and often require deeper understanding of the subject matter.

\input{up2020-en-1.tex}
\input{up2020-en-2.tex}
\input{up2020-en-3.tex}

\newpart
\printbibliography[heading = bibintoc, title = References]
\newpage
\specialsec{Afterword}\label{sec:afterword}

Soon after beginning to learn Linux, I bought the book \emph{Classic Shell
Scripting}, which initiated my first enthusiastic experience in formal
programming.  My interest was exactly because of the word frequency program
mentioned in Section~\ref{sec:shell}, used by the book as an example for the
power of the Unix philosophy.  (Note how this is similar to V.~I.\ Arnold's
reaction to L.~A.\ Tumarkin's teaching practice -- \cf~Section~%
\ref{sec:cognitive}.)  Although the Unix philosophy (under the name
``Software Tools philosophy'') was followed throughout the book, I did
not intentionally follow the philosophy until I attended the \emph{Linux
Programming Environment} course in Autumn 2009 by Mr.\ Dong-Gang Cao
at my university, where the notion was formally given and emphasised.

My life with Linux was relatively peaceful, until systemd came into the picture:
in 2011 and 2012, the developers of systemd pushed for the addition of logind
as a dependency of GNOME 3\cupercite{poettering2011a}, and merged udev into
systemd\cupercite{sievers2012}; vehement controversies and flamewars beginning
from then only began to wane in the last two years\cupercite{slashdot:systemd},
and it is my explicit wish that this document contributes to the demise of
systemd and the proprietary practices associated with it.  ``Every cloud has
a silver lining'': in the search for a systemd substitute I found the s6 project
in Summer 2014, and studying the design and implementation of software in the
style by Daniel J.\ Bernstein proved to be an invaluable practice in application
of the Unix philosophy.  During the process, the questions ``is the Unix
philosophy now outdated?'' and ``what is the social value of the Unix
philosophy?'' became recurring issues in my mind, which resulted
in the first two main parts of this document.

In Spring 2018, I envisioned a somehow C-like language based on S-expressions
when considering the inexpressiveness of C\cupercite{vector2018b}, roughly
one year after stumbling upon scsh\cupercite{angelbeats2018}, although I
became aware of the elegance of Lisp perhaps as early as 2012, mainly from
the writings by Yin Wang (while he is highly controversial in the Chinese
open-source community, he often makes very interesting points).  The
inspiration for said C-like language can actually be traced earlier, when
I began thinking about the possibility of a shell-like language that also
supports Bernstein chainloading\cupercite{vector2016a} in the beginning of
2016.  I finally came up with a practical minimalist plan for such a language%
\cupercite{vector2018c} in Spring 2018, and meanwhile came across the idea of
``Lisp~/ C reconciliation'', which grew into the third part of this document
through a discussion thread\cupercite{stevel2018} in the Gentoo forums.

Before ending this document, I would like to thank the Linux Club of Peking
University and the skarnet software community, without which I would probably
be yet another highly amateurish dabbler in Linux, and this document would
have never come into existence.  I would also like to thank the general
open-source community, including the Unix circle, the Lisp circle and more,
encompassing people pro-systemd and anti-systemd, for providing rich and
diverse technical materials and opinions, which is a priceless offering.
Inspired by the musical settings in \emph{Thunderstorm} by Yu Cao,
and the appendices to \emph{The \hologo{TeX}book} and \emph{The
\hologo{METAFONT}book} by Donald Knuth, the structure of this
document was modeled after the \emph{Mass in B Minor} by J.~S.\ Bach.
\end{document}

